---
title: "DengAI_EDA"
author: "Oren Jalon"
date: "November 19, 2018"
output: html_document
---
# DENGAI 

The following code is my submission to the DengAI competition on DrivenData.org.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# DATA PREPARATION

Let's begin by putting in the libraries we need and importing the data

## LIBRARIES
```{r library, echo = TRUE}
library(tidyverse)
library(caret)
library(reshape2)
```
## IMPORT DATA

Make sure that the files are in the same working directory as this RMD file (-- use getwd() to check --)

```{r inport, echo = TRUE, message = FALSE}

d_train <- read_csv(file = "dengue_features_train.csv")
d_labels <- read_csv(file = "dengue_labels_train.csv")
d_test <- read_csv(file = "dengue_features_train.csv")
```
### Rescale the variables so that it is all in Celcius and mm
```{r rescaling, echo=TRUE}
d_train$reanalysis_dew_point_temp_k <- d_train$reanalysis_dew_point_temp_k - 273.15
d_test$reanalysis_dew_point_temp_k <- d_test$reanalysis_dew_point_temp_k - 273.15

d_train$reanalysis_air_temp_k <- d_train$reanalysis_air_temp_k - 273.15
d_test$reanalysis_air_temp_k <- d_test$reanalysis_air_temp_k - 273.15

d_train$reanalysis_max_air_temp_k <- d_train$reanalysis_max_air_temp_k - 273.15
d_test$reanalysis_max_air_temp_k <- d_test$reanalysis_max_air_temp_k - 273.15

d_train$reanalysis_min_air_temp_k <- d_train$reanalysis_min_air_temp_k - 273.15
d_test$reanalysis_min_air_temp_k <- d_test$reanalysis_min_air_temp_k - 273.15

d_train$reanalysis_avg_temp_k <- d_train$reanalysis_avg_temp_k - 273.15
d_test$reanalysis_avg_temp_k <- d_test$reanalysis_avg_temp_k - 273.15

#!!!tdtr does not appear to be in Kelvin
# d_train$reanalysis_tdtr_k <- d_train$reanalysis_tdtr_k - 273.15
# d_test$reanalysis_tdtr_k <- d_test$reanalysis_tdtr_k - 273.15

summary(d_train$reanalysis_dew_point_temp_k)
summary(d_train$reanalysis_air_temp_k)
summary(d_train$reanalysis_max_air_temp_k)
summary(d_train$reanalysis_min_air_temp_k)
summary(d_train$reanalysis_avg_temp_k)
summary(d_train$reanalysis_tdtr_k)
```
We seem to have some missing values but we will investigate this more later.

Now let's merge the datasets (train and labels) so that we can have the the features and the classifier in the same place.  We will also split the dataset by city.  Finally, we will merge the test and training set so that we can see what all the climate data looks like across the entire recorded period.  This is just for exploration and will not be used for our analysis later.

## MERGE TRAIN & LABELS
```{r add labels, echo=TRUE}
df <- merge(d_train, d_labels, by=c("city","year","weekofyear"))

```
## FILTER BY CITY (SJ, IQ)
```{r sj & iq test, echo=TRUE}

sj <- df[df$city == "sj",] 
iq <- df[df$city == "iq",] 
```


# INITIAL & EXPLORATORY ANALYSIS

Let's review the summary statistics for the dataset.  We will use the skimr library to get a cleaner view of the data.


## Summary Stats
```{r summary, echo=TRUE}
library(skimr)

skimmed.sj <- skim_to_wide(sj[,c(-1,-4)])
skimmed.sj

skimmed.iq <- skim_to_wide(iq[,c(-1,-4)])
skimmed.iq

rm(skimmed.sj, skimmed.iq)

```

We can see that SJ has 936 weeks with a mean of 34.2 cases per week.  IQ has 520 weeks with a mean of 7.6 weekly cases.  There are no missing values for our total cases classifer but there are climate data features which have many missing values.  

In this section, we summary the value of the data frames (together and by city).  We also create the following graphs

1.  Wilcoxon test for test of means between cities
2. Frequency histograms
3. Bivariate analysis - line graphs for time analysis
4. Bivariate analysis - box plot by week for determining seasonality
5. Bivariate analysis - total cases by time and by week
6. Bivariate analysis - scatterplot for total_cases by other variables

## Compare the means between same variables in different cities

We need to know if the two cities should be analysed together or separately.  Let's do a wilcoxon test (non-parametric) to compare the means of the same climate features between the two different cities. 

```{r means between cities, echo = TRUE}
cnames <- colnames(sj)
for (i in 5:(ncol(sj))){
  wilt <- wilcox.test(sj[,i],iq[,i])
  print(cnames[i])
  print(wilt)
}

rm(cnames, i, wilt)
```
We can see that the same feature is significantly different in each city (look at the p-values) 

## GRAPH: Frequency histogram of all variables in training set for **SJ**

Let's review the distribution of each of the variables.  Some appear to be normally distributed while others are not.

```{r graph histogram SJ, echo=TRUE}
cnames <- colnames(sj) 
par(mfrow=c(1,2))
for (i in 5:ncol(sj)) {
  hist(sj[,i], 
      breaks = 20,
      xlab = cnames[i], 
      main = paste(cnames[i], sep = ": "))
}

rm(cnames, i)

```

## GRAPH: Frequency histogram of all variables in training set for **IQ**

Same as above but only for IQ.

```{r graph histogram IQ, echo=TRUE}
cnames <- colnames(df) 
par(mfrow=c(1,2))
for (i in 5:(ncol(df))) {
 hist(df[df$city == "iq",i],
      breaks = 20,
      xlab = cnames[i],
      main = paste("Freq Histogram for IQ", cnames[i], sep = ": "))
}

rm(cnames, i)

```

## GRAPH: Climate variables by time for **SJ**

Here we are reviewing the data by time for SJ
  

```{r graph sj climate by time, echo=TRUE}
cnames <- colnames(sj) 
par(mfrow=c(2,2))
for (i in 5:(ncol(sj))) {
  gg1 <- ggplot(sj,
                aes(x=week_start_date, 
                    y = sj[,i])) +
    geom_line() +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)


```

## GRAPH: Climate variables by time for **IQ**

Here we are reviewing the data by time for IQ

```{r graph iq climate by time, echo=TRUE}

cnames <- colnames(iq) 
par(mfrow=c(2,2))
for (i in 5:(ncol(iq))) {
  gg1 <- ggplot(iq,
                aes(x=week_start_date, 
                    y = iq[,i])) +
    geom_line() +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

## GRAPH: Climate variables by week for **SJ**

Now let's review the climate variables by week to determine seasonality of the data (SJ ONLY)

```{r graph sj by week, echo=TRUE}
library(ggplot2)

cnames <- colnames(sj) 
par(mfrow=c(2,2))
for (i in 5:(ncol(sj))) {
  gg1 <- ggplot(sj,
                aes(x=weekofyear, 
                    y = sj[,i], 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,2)) +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

## GRAPH: Climate variables by week for **IQ**

Now let's review the climate variables by week to determine seasonality of the data (IQ ONLY)

```{r graph iq by week, echo=TRUE}
library(ggplot2)

cnames <- colnames(iq) 
par(mfrow=c(2,2))
for (i in 5:(ncol(iq))) {
  gg1 <- ggplot(iq,
                aes(x=weekofyear, 
                    y = iq[,i], 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,2)) +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

## GRAPH: Total_cases by time for **SJ**, **IQ**

Same as above but a little more descriptive

```{r graph total cases time analysis, echo=TRUE}
library(ggplot2)

par(mfcol=c(1,3))

#Dengue Cases for San Jose
ggplot(data = df[df$city == "sj",], aes(x=week_start_date, y=total_cases)) +
       geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Total Dengue Cases in San Jose",
       subtitle = paste(min(df$week_start_date[df$city == "sj"]),max(df$week_start_date[df$city == "sj"]), sep = " to "),
       x = "Date", y = "Total dengue cases")

# Dengue Cases for Iquitos
ggplot(data = df[df$city == "iq",], aes(x=week_start_date, y=total_cases)) +
       geom_bar(stat = "identity", fill = "green") +
  labs(title = "Total Dengue Cases in Iquitos",
       subtitle = paste(min(df$week_start_date[df$city == "iq"]),max(df$week_start_date[df$city == "iq"]), sep = " to "),
       x = "Date", y = "Total dengue cases")

```

## GRAPH: Average Total_cases by week for **SJ**, **IQ**
Line graph of all data by total cases.  This uses only the training set.
```{r graph total cases by week, echo=TRUE}
library(ggplot2)

gg1 <- ggplot(sj,
                aes(x=weekofyear, 
                    y = total_cases, 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,1)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") +
    ylab("Total cases") +
    ggtitle(paste("Boxplot: Total cases by Week for SJ")) 

    print(gg1)
    
gg3 <- ggplot(data=sj, aes(x=weekofyear, y=total_cases)) +
  geom_bar(stat="summary", fun.y = "mean") +
  ggtitle(paste("Bar graph: Average total cases by Week for SJ")) +
  scale_x_continuous(breaks = seq(1,52, 2))

print(gg3)

    
gg2 <- ggplot(iq,
                aes(x=weekofyear, 
                    y = total_cases, 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,1)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") +
    ylab("Total cases") +
    ggtitle(paste("Boxplot: Total cases by Week for IQ")) 

    print(gg2)

gg4 <- ggplot(data=iq, aes(x=weekofyear, y=total_cases)) +
  geom_bar(stat="summary", fun.y = "mean") +
  ggtitle(paste("Bar graph: Average total cases by Week for IQ")) +
  scale_x_continuous(breaks = seq(1,52, 2))

print(gg4)
    
    rm(gg1, gg2, gg3, gg4)


```

## GRAPH: Total_cases by climate variables for **SJ**
Same as above but for SJ
```{r graph SJ total cases analysis, echo=TRUE}


cnames <- colnames(df) 
par(mfrow=c(2,2))
for (i in 5:(ncol(df)-1)) {
 plot(df$total_cases[df$city == "sj"],
      df[df$city == "sj",i], 
      cex = 0.5, 
      pch = 19,
      ylim = c(min(df[,i],na.rm=TRUE), max(df[,i],na.rm=TRUE)),
      main = paste("Total_cases for SJ by climate variables", cnames[i], sep = ": "),
      ylab = cnames[i])
 
}

rm(cnames, i)

```

## GRAPH: Total_cases by climate variables for **IQ**
Same as above but for IQ.
```{r graph IQ total cases analysis, echo=TRUE}


cnames <- colnames(df) 
par(mfrow=c(2,2))
for (i in 5:(ncol(df)-1)) {
 plot(df$total_cases[df$city == "iq"],
      df[df$city == "iq",i], 
      cex = 0.5, 
      pch = 19,
      ylim = c(min(df[,i],na.rm=TRUE), max(df[,i],na.rm=TRUE)),
      main = paste("Total_cases for IQ by climate variables", cnames[i], sep = ": "),
      ylab = cnames[i])
 
}

rm(cnames, i)

```


# ANALYSIS OF OUTLIERS
### GRAPH: Boxplot of climate variables (test and train)
Boxplot includes test and training set - NA still included

```{r graph boxplot, echo=TRUE}
library(ggplot2)
cnames <- colnames(df) 
for (i in 5:(ncol(df))) {
 p <- ggplot(df, aes(x=city, y = df[,i], fill = city)) + 
  geom_boxplot() +
   labs(title = "Boxplot of climate variables",
       subtitle = cnames[i],
       x = "City", y = cnames[i])
 print(p)
}
rm(cnames, i, p)
```





























