---
title: "DengAI_EDA"
author: "Oren Jalon"
date: "November 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#DATA PREPARATION
##LIBRARIES
```{r library, echo = TRUE}
library(tidyverse)
library(caret)
library(reshape2)
```
##IMPORT DATA
```{r inport, echo = TRUE, message = FALSE}

d_train <- read_csv("D:/Google Drive/RYERSON/CKME 136/DengAI/DATASET/dengue_features_train.csv")
d_labels <- read_csv("D:/Google Drive/RYERSON/CKME 136/DengAI/DATASET/dengue_labels_train.csv")
d_test <- read_csv("D:/Google Drive/RYERSON/CKME 136/DengAI/DATASET/dengue_features_train.csv")
```
###Rescale the variables so that it is all in Celcius and mm
```{r rescaling, echo=TRUE}
d_train$reanalysis_dew_point_temp_k <- d_train$reanalysis_dew_point_temp_k - 273.15
d_test$reanalysis_dew_point_temp_k <- d_test$reanalysis_dew_point_temp_k - 273.15

d_train$reanalysis_air_temp_k <- d_train$reanalysis_air_temp_k - 273.15
d_test$reanalysis_air_temp_k <- d_test$reanalysis_air_temp_k - 273.15

d_train$reanalysis_max_air_temp_k <- d_train$reanalysis_max_air_temp_k - 273.15
d_test$reanalysis_max_air_temp_k <- d_test$reanalysis_max_air_temp_k - 273.15

d_train$reanalysis_min_air_temp_k <- d_train$reanalysis_min_air_temp_k - 273.15
d_test$reanalysis_min_air_temp_k <- d_test$reanalysis_min_air_temp_k - 273.15

d_train$reanalysis_avg_temp_k <- d_train$reanalysis_avg_temp_k - 273.15
d_test$reanalysis_avg_temp_k <- d_test$reanalysis_avg_temp_k - 273.15

#!!!tdtr does not appear to be in Kelvin
# d_train$reanalysis_tdtr_k <- d_train$reanalysis_tdtr_k - 273.15
# d_test$reanalysis_tdtr_k <- d_test$reanalysis_tdtr_k - 273.15

summary(d_train$reanalysis_dew_point_temp_k)
summary(d_train$reanalysis_air_temp_k)
summary(d_train$reanalysis_max_air_temp_k)
summary(d_train$reanalysis_min_air_temp_k)
summary(d_train$reanalysis_avg_temp_k)
summary(d_train$reanalysis_tdtr_k)
```


##MERGE TRAIN & LABELS
```{r add labels, echo=TRUE}
df <- merge(d_train, d_labels, by=c("city","year","weekofyear"))

```
###FILTER BY CITY (SJ, IQ)
```{r sj & iq test, echo=TRUE}

sj <- df[df$city == "sj",] 
iq <- df[df$city == "iq",] 
```
##MERGE TRAIN & TEST
Merge test and train set without the total_cases
```{r full, echo=TRUE}
df_all <- rbind(d_train,d_test)
```
###FILTER BY CITY (SJ_ALL, IQ_ALL)
```{r sj & iq testall, echo=TRUE}
iq_all <- df_all[df_all$city == 'iq', ]
sj_all <- df_all[df_all$city == 'sj', ]
```

#PRELIMINARY DATA PREPARATION

This section includes importing the data, creating of new variables and establishing the dataframes for the initial analysis

##Summary Stats
```{r summary, echo=TRUE}
library(skimr)

skimmed.sj <- skim_to_wide(sj[,c(-1,-4)])
skimmed.sj

skimmed.iq <- skim_to_wide(iq[,c(-1,-4)])
skimmed.iq

skimmed.sj_all <- skim_to_wide(sj_all[,c(-1,-4)])
skimmed.sj_all

skimmed.iq_all <- skim_to_wide(iq_all[,c(-1,-4)])
skimmed.iq_all

rm(skimmed.sj, skimmed.iq, skimmed.iq_all, skimmed.sj_all)

```
#DATAFRAME CLEANUP 1
Clean up all the extra dataframes produced during the exploratory analysis

```{r df cleanup 1}
#rm(d_test,
   # d_train,
   # dengue_labels_train,
   # sj_test,
   # sj_features_train,
   # sj_labels_train,
   # iq_test,
   # iq_features_train,
   # iq_labels_train,
   # df,
   # iq,
   # sj,
   # df,
   # submission_format
   #  )



```
#INITIAL & EXPLORATORY ANALYSIS

In this section, we summary the value of the data frames (together and by city).  We also create the following graphs

1. Frequency histograms
2. Bivariate analysis - line graphs for time analysis
3. Bivariate analysis - scatterplot for total_cases by other variables
4. Wilcoxon test for test of means between cities

###Compare the means between same variables in different cities
We can see that the same feature is significantly different in each city 

```{r means between cities, echo = TRUE}
cnames <- colnames(sj)
for (i in 5:(ncol(sj))){
  wilt <- wilcox.test(sj[,i],iq[,i])
  print(cnames[i])
  print(wilt)
}

rm(cnames, i, wilt)
```

###GRAPH: Frequency histogram of all variables in training set for **SJ**
Same as above but only for SJ.
```{r graph histogram SJ, echo=TRUE}
cnames <- colnames(sj) 
par(mfrow=c(1,2))
for (i in 5:ncol(sj)) {
  hist(sj[,i], 
      breaks = 20,
      xlab = cnames[i], 
      main = paste(cnames[i], sep = ": "))
}

rm(cnames, i)

```

###GRAPH: Frequency histogram of all variables in training set for **IQ**
Same as above but only for IQ.
```{r graph histogram IQ, echo=TRUE}
cnames <- colnames(df) 
par(mfrow=c(1,2))
for (i in 5:(ncol(df))) {
 hist(df[df$city == "iq",i],
      breaks = 20,
      xlab = cnames[i],
      main = paste("Freq Histogram for IQ", cnames[i], sep = ": "))
}

rm(cnames, i)

```

###GRAPH: Climate variables by time for **SJ**
Includes all the data from test and training set by time for SJ therefore the total_cases in not included.  Total_cases by time is done separately.

```{r graph sj climate by time, echo=TRUE}
cnames <- colnames(sj_all) 
par(mfrow=c(2,2))
for (i in 5:(ncol(sj_all))) {
  gg1 <- ggplot(sj_all,
                aes(x=week_start_date, 
                    y = sj_all[,i])) +
    geom_line() +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)
```

###GRAPH: Climate variables by time for **IQ**
Includes all the data from test and training set by time for I therefore the total_cases in not included.  Total_cases by time is done separately.
```{r graph iq climate by time, echo=TRUE}

cnames <- colnames(iq_all) 
par(mfrow=c(2,2))
for (i in 5:(ncol(iq_all))) {
  gg1 <- ggplot(iq_all,
                aes(x=week_start_date, 
                    y = iq_all[,i])) +
    geom_line() +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

###GRAPH: Climate variables by week for **SJ**
Includes all the data from test and training set by time for SJ therefore the total_cases in not included.  Total_cases by time is done separately.

```{r graph sj by week, echo=TRUE}
library(ggplot2)

cnames <- colnames(sj_all) 
par(mfrow=c(2,2))
for (i in 5:(ncol(sj_all))) {
  gg1 <- ggplot(sj,
                aes(x=weekofyear, 
                    y = sj_all[,i], 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,2)) +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

###GRAPH: Climate variables by week for **IQ**
Includes all the data from test and training set by time for I therefore the total_cases in not included.  Total_cases by time is done separately.
```{r graph iq by week, echo=TRUE}
library(ggplot2)

cnames <- colnames(iq_all) 
par(mfrow=c(2,2))
for (i in 5:(ncol(iq_all))) {
  gg1 <- ggplot(iq_all,
                aes(x=weekofyear, 
                    y = iq_all[,i], 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,2)) +
    ylab(cnames[i]) +
    ggtitle(paste(cnames[i])) 

    print(gg1)
  }

rm(cnames, i, gg1)

```

###GRAPH: Total_cases by time for **SJ**, **IQ**
Line graph of all data by total cases.  This uses only the training set.
```{r graph total cases time analysis, echo=TRUE}
library(ggplot2)

par(mfcol=c(1,3))

#Dengue Cases for San Jose
ggplot(data = df[df$city == "sj",], aes(x=week_start_date, y=total_cases)) +
       geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Total Dengue Cases in San Jose",
       subtitle = paste(min(df$week_start_date[df$city == "sj"]),max(df$week_start_date[df$city == "sj"]), sep = " to "),
       x = "Date", y = "Total dengue cases")

# Dengue Cases for Iquitos
ggplot(data = df[df$city == "iq",], aes(x=week_start_date, y=total_cases)) +
       geom_bar(stat = "identity", fill = "green") +
  labs(title = "Total Dengue Cases in Iquitos",
       subtitle = paste(min(df$week_start_date[df$city == "iq"]),max(df$week_start_date[df$city == "iq"]), sep = " to "),
       x = "Date", y = "Total dengue cases")

```

###GRAPH: Average Total_cases by week for **SJ**, **IQ**
Line graph of all data by total cases.  This uses only the training set.
```{r graph total cases by week, echo=TRUE}
library(ggplot2)

gg1 <- ggplot(sj,
                aes(x=weekofyear, 
                    y = total_cases, 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,1)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") +
    ylab("Total cases") +
    ggtitle(paste("Boxplot: Total cases by Week for SJ")) 

    print(gg1)
    
gg3 <- ggplot(data=sj, aes(x=weekofyear, y=total_cases)) +
  geom_bar(stat="summary", fun.y = "mean") +
  ggtitle(paste("Bar graph: Average total cases by Week for SJ")) +
  scale_x_continuous(breaks = seq(1,52, 2))

print(gg3)

    
gg2 <- ggplot(iq,
                aes(x=weekofyear, 
                    y = total_cases, 
                    group = weekofyear)) +
    geom_boxplot() +
    scale_x_continuous(breaks=seq(1,52,1)) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") +
    ylab("Total cases") +
    ggtitle(paste("Boxplot: Total cases by Week for IQ")) 

    print(gg2)

gg4 <- ggplot(data=iq, aes(x=weekofyear, y=total_cases)) +
  geom_bar(stat="summary", fun.y = "mean") +
  ggtitle(paste("Bar graph: Average total cases by Week for IQ")) +
  scale_x_continuous(breaks = seq(1,52, 2))

print(gg4)
    
    rm(gg1, gg2, gg3, gg4)


```

###GRAPH: Total_cases by climate variables for **SJ**
Same as above but for SJ
```{r graph SJ total cases analysis, echo=TRUE}


cnames <- colnames(df) 
par(mfrow=c(2,2))
for (i in 5:(ncol(df)-1)) {
 plot(df$total_cases[df$city == "sj"],
      df[df$city == "sj",i], 
      cex = 0.5, 
      pch = 19,
      ylim = c(min(df[,i],na.rm=TRUE), max(df[,i],na.rm=TRUE)),
      main = paste("Total_cases for SJ by climate variables", cnames[i], sep = ": "),
      ylab = cnames[i])
 
}

rm(cnames, i)

```

###GRAPH: Total_cases by climate variables for **IQ**
Same as above but for IQ.
```{r graph IQ total cases analysis, echo=TRUE}


cnames <- colnames(df) 
par(mfrow=c(2,2))
for (i in 5:(ncol(df)-1)) {
 plot(df$total_cases[df$city == "iq"],
      df[df$city == "iq",i], 
      cex = 0.5, 
      pch = 19,
      ylim = c(min(df[,i],na.rm=TRUE), max(df[,i],na.rm=TRUE)),
      main = paste("Total_cases for IQ by climate variables", cnames[i], sep = ": "),
      ylab = cnames[i])
 
}

rm(cnames, i)

```


##Compare similar variable values within the dataset
There are several variables which appear to be the same feature but taken from a different source.  For example, station_precip_mm and	precipitation_amt_mm and reanalysis_sat_precip_amt_mm all appear to be the same "Total Precipitation value"  Only one should be kept if they are the same.

###Difference in max air temp
"station_max_temp_c"" and "reanalysis_max_air_temp_k" (scaled to Celcius)
```{r max air, echo= TRUE}
library(ggplot2)

#generate a difference in max temp variable
sj$max_air_diff <- sj$station_max_temp_c - sj$reanalysis_max_air_temp_k

#barplot the difference by year
ggplot(sj,aes(x=year, y=max_air_diff))+
  geom_bar(stat='identity')

#box plot difference by year
ggplot(sj, aes(x=year, y = max_air_diff, group = year)) +   geom_boxplot() 

#Add month to the dataframe
sj$month <- as.POSIXlt(sj$week_start_date)$mon +1

#box plot difference by month
ggplot(sj, aes(x=month, y = max_air_diff, group = month)) +   geom_boxplot() + scale_x_continuous(breaks=seq(1,12,1))

sj$max_air_diff <- NULL
sj$month <- NULL
```

###Difference in min air temp
"station_min_temp_c"" and "reanalysis_min_air_temp_k" (scaled to Celcius)
```{r min air, echo= TRUE}
library(ggplot2)

#generate a difference in max temp variable
sj$min_air_diff <- sj$station_min_temp_c - sj$reanalysis_min_air_temp_k

#barplot the difference by year
ggplot(sj,aes(x=year, y=min_air_diff))+
  geom_bar(stat='identity')

#box plot difference by year
ggplot(sj, aes(x=year, y = min_air_diff, group = year)) +   geom_boxplot() 

#Add month to the dataframe
sj$month <- as.POSIXlt(sj$week_start_date)$mon +1

#box plot difference by month
ggplot(sj, aes(x=month, y = min_air_diff, group = month)) +   geom_boxplot() + scale_x_continuous(breaks=seq(1,12,1))

sj$min_air_diff <- NULL
sj$month <- NULL
```

###Difference in average air temp
"station_avg_temp_c"" and "reanalysis_avg_temp_k" (scaled to Celcius)
```{r avg air, echo= TRUE}
library(ggplot2)

#generate a difference in max temp variable
sj$avg_air_diff <- sj$station_avg_temp_c - sj$reanalysis_avg_temp_k

#barplot the difference by year
ggplot(sj,aes(x=year, y=avg_air_diff))+
  geom_bar(stat='identity')

#box plot difference by year
ggplot(sj, aes(x=year, y = avg_air_diff, group = year)) +   geom_boxplot() 

#Add month to the dataframe
sj$month <- as.POSIXlt(sj$week_start_date)$mon +1

#box plot difference by month
ggplot(sj, aes(x=month, y = avg_air_diff, group = month)) +   geom_boxplot() + scale_x_continuous(breaks=seq(1,12,1))

sj$avg_air_diff <- NULL
sj$month <- NULL
```

###Difference in total precipitation
"station_precip_mm", "precipitation_amt_mm", "reanalysis_sat_precip_amt_mm", "reanalysis_precip_amt_kg_per_m2" 

```{r total precip, echo= TRUE}
library(ggplot2)

precip <- c("station_precip_mm", "precipitation_amt_mm", "reanalysis_sat_precip_amt_mm", "reanalysis_precip_amt_kg_per_m2")

#Add month to the dataframe
sj$month <- as.POSIXlt(sj$week_start_date)$mon +1



for (i in 1:3){
  par(mfrow=c(1,3))
  #generate the first variable in the list
  p1 <- precip[i]
  ind1 <- which(colnames(sj)==p1)
  for (j in ((i+1):4)){
    #generate the next variable in the list
    p2 <- precip[j]
  ind2 <- which(colnames(sj)==p2)
  #generate a difference variable 
   sj$diff <- sj[,ind1] - sj[,ind2]
   
   #barplot the difference by year
   gg1 <-ggplot(sj,
                 aes(x=year, y=diff))+
      geom_bar(stat = "identity", fill="steelblue") + 
      ggtitle(paste(p1, "&", p2))
    print(gg1)
    
    #box plot the difference by year
   gg2 <-ggplot(sj,
                 aes(x=year, y=diff, group = year)) +
      geom_boxplot() + 
      ggtitle(paste(p1, "&", p2))
    print(gg2)
    
    #box plot difference by month
    gg3 <- ggplot(sj, 
                  aes(x=month, y = diff, group = month)) +
      geom_boxplot() +
      scale_x_continuous(breaks=seq(1,12,1)) +
      ggtitle(paste(p1, "&", p2))
    print(gg3)
  }
}

sj$diff <- NULL
sj$month <- NULL 

rm(gg1, gg2, gg3, i, ind1, ind2, j, p1, p2, precip)

```
#ANALYSIS OF OUTLIERS
###GRAPH: Boxplot of climate variables (test and train)
Boxplot includes test and training set - NA still included

```{r graph boxplot, echo=TRUE}
library(ggplot2)
cnames <- colnames(df) 
for (i in 5:(ncol(df))) {
 p <- ggplot(df, aes(x=city, y = df[,i], fill = city)) + 
  geom_boxplot() +
   labs(title = "Boxplot of climate variables",
       subtitle = cnames[i],
       x = "City", y = cnames[i])
 print(p)
}
rm(cnames, i, p)
```

###GRAPH: Boxplot of total cases
```{r graph boxplot total cases, echo=TRUE}
library(ggplot2)
ggplot(df, aes(x=city, y = total_cases, fill = city)) + 
  geom_boxplot() +
   labs(title = "Boxplot of Total_cases",
       x = "City", y = "Total_cases")
```






























